---
title: 'Data Science 3: Assignment 2'
author: "Benedek Pásztor"
date: '2019 04 06 '
output: html_document
---

## Deep neural nets with `keras`

The [homepage](https://keras.rstudio.com/) has great descrpitions, expamples
and tutorials. Cheatsheet [here](https://www.rstudio.com/resources/cheatsheets/). 

```{r}
# install.packages("keras")
library(keras)
# install_keras()
```

  
```{r}
library(keras)
library(here)
library(grid)
library(magick)  # not absolutely necessary
```


```{r}
fashion_mnist <- dataset_fashion_mnist()
x_train <- fashion_mnist$train$x
y_train <- fashion_mnist$train$y
x_test <- fashion_mnist$test$x
y_test <- fashion_mnist$test$y
```


### a. Show some example images from the data.

```{r, fig.width=2, fig.height=2}
show_mnist_image <- function(x) {
  image(1:28, 1:28, t(x)[,nrow(x):1],col=gray((0:255)/255)) 
}

show_mnist_image(x_train[18, , ])
```


```{r, fig.width=2, fig.height=2}
show_mnist_image(x_train[27, , ])
```


```{r, fig.width=2, fig.height=2}
show_mnist_image(x_train[98, , ])
```


### b. Train a fully connected deep network to predict items.

##### - Normalize the data similarly to what we saw with MNIST.

```{r}
# reshape
x_train <- array_reshape(x_train, c(dim(x_train)[1], 784)) 
x_test <- array_reshape(x_test, c(dim(x_test)[1], 784)) 
# rescale
x_train <- x_train / 255
x_test <- x_test / 255

# one-hot encoding of the target variable
y_train <- to_categorical(y_train, 10)
y_test <- to_categorical(y_test, 10)
```

##### - Experiment with network architectures and settings (number of hidden layers, number of nodes, activation functions, dropout, etc.)

###### Model 1

```{r}
model <- keras_model_sequential() 
model <- model %>% 
  layer_dense(units = 128, activation = 'relu', input_shape = c(784)) %>%
  layer_dropout(rate = 0.3) %>%
  layer_dense(units = 10, activation = 'softmax')
```

```{r}
summary(model)
```
```{r}
model %>% compile(
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)
```

```{r message = FALSE, tidy = TRUE}
history <- model %>% fit(
  x_train, y_train, 
  epochs = 50, batch_size = 128, 
  validation_split = 0.2
)
```

```{r}
model %>% evaluate(x_test, y_test)
```

###### Model 2

##### - Explain what you have tried, what worked and what did not. Present a final model.


##### - Make sure that you use enough epochs so that the validation error starts flattening out - provide a plot about the training history (plot(history))

```{r}
plot(history)
```



###### Error check


```{r}
predicted_classes_test <- model %>% predict_classes(x_test)
real_classes_test <- as.numeric(fashion_mnist$test$y)
library(data.table)
dt_pred_vs_real <- data.table(predicted = predicted_classes_test, real = real_classes_test)

library(ggplot2)
ggplot(dt_pred_vs_real[, .N, by = .(predicted, real)], aes(predicted, real)) +
  geom_tile(aes(fill = N), colour = "white") +
  scale_x_continuous(breaks = 0:9) +
  scale_y_continuous(breaks = 0:9) +
  geom_text(aes(label = sprintf("%1.0f", N)), vjust = 1, color = "white") +
  scale_fill_viridis_c() +
  theme_bw() + theme(legend.position = "none")
```

```{r}
dt_pred_vs_real[, row_number := 1:.N]
indices_of_mistakes <- dt_pred_vs_real[predicted != real][["row_number"]]
```


```{r, fig.width=2, fig.height=2}
ix <- indices_of_mistakes[1]

dt_pred_vs_real[row_number == ix]
show_mnist_image(fashion_mnist$test$x[ix, , ])
```

```{r, fig.width=2, fig.height=2}
ix <- indices_of_mistakes[11]

dt_pred_vs_real[row_number == ix]
show_mnist_image(fashion_mnist$test$x[ix, , ])
```


##### - Evaluate the model on the test set. How does test error compare to validation error?


### c. Try building a convolutional neural network and see if you can improve test set performance.


```{r}

fashion_mnist <- dataset_fashion_mnist()
x_train <- fashion_mnist$train$x
y_train <- fashion_mnist$train$y
x_test <- fashion_mnist$test$x
y_test <- fashion_mnist$test$y

x_train <- array_reshape(x_train, c(nrow(x_train), 28, 28, 1))
x_test <- array_reshape(x_test, c(nrow(x_test), 28, 28, 1))

# rescale
x_train <- x_train / 255
x_test <- x_test / 255

# one-hot encoding of the target variable
y_train <- to_categorical(y_train, 10)
y_test <- to_categorical(y_test, 10)
```


```{r}
cnn_model <- keras_model_sequential() 
cnn_model %>% 
  layer_conv_2d(filters = 32,
                kernel_size = c(3, 3), 
                activation = 'relu',
                input_shape = c(28, 28, 1)) %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_dropout(rate = 0.25) %>%
  layer_flatten() %>% 
  layer_dense(units = 32, activation = 'relu') %>% 
  layer_dense(units = 10, activation = 'softmax')
```

```{r}
summary(cnn_model)
```



```{r}
cnn_model %>% compile(
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)
```

```{r}
history <- cnn_model %>% fit(
  x_train, y_train, 
  epochs = 2, batch_size = 128, 
  validation_split = 0.2
)
```

```{r}
cnn_model %>% evaluate(x_test, y_test)
```



### d. Just like before, experiment with different network architectures, regularization techniques and present your findings



# 2. Hot dog or not hot dog?



### a. Pre-process data so that it is acceptable by Keras (set folder structure, bring images to the same size, etc).

### b. Estimate a convolutional neural network to predict if an image contains a hot dog or not. Evaluate your model on the test set.

### c. Could data augmentation techniques help with achieving higher predictive accuracy? Try some augmentations that you think make sense and compare

### d. Try to rely on some pre-built neural networks to aid prediction. Can you achieve a better performance using transfer learning for this problem?